---
title: Wine Quality Analysis

# Use letters for affiliations
author:
  - name: Hirley Dayan Lourenço da Silva *and* Marcia Maria Parmigiani Martins
    affiliation:
# address:
#  - code: A
#    address: Institute of Smoke and Magic, University of Sometown, Sometown, XY, 12345
#  - code: A
#    address: Department of Neat Tricks, Whereever State University, Someplace, MC, 67890
    
# Optional: line of arbitrary text with additional information.
# Could be used, for example, to mention the bibliographic info in a post-print.
# If not specified, defaults to "This version was compiled on \today"
date_subtitle: "Unicamp Data Science - Supervised Learning Module "

# For footer text  TODO(fold into template, allow free form two-authors)
# lead_author_surname: Author and Author

# Place eg a DOI URL or CRAN Package URL here
# doi_footer: "https://cran.r-project.org/package=YourPackage"

# Abstract
abstract: |
  Logistic regression is used to describe data and to explain the relationship 
  between one dependent binary variable and one or more independent variables. 
  This work explores Logistic Regression techniques for creating a model for 
  classifying the quality of wines based on a few chemical features. 

# Optional: Acknowledgements
# acknowledgements: |
#  This template package builds upon, and extends, the work of the excellent
#  [rticles](https://cran.r-project.org/package=rticles) package, and both packages rely on the
#  [PNAS LaTeX](http://www.pnas.org/site/authors/latex.xhtml) macros. Both these sources are
#  gratefully acknowledged as this work would not have been possible without them.  Our extensions
#  are under the same respective licensing term
#  ([GPL-3](https://www.gnu.org/licenses/gpl-3.0.en.html) and
#  [LPPL (>= 1.3)](https://www.latex-project.org/lppl/)).

# Optional: One or more keywords
keywords:
  - Logistic Regression
  - glm
  - glmnet

# Paper size for the document, values of letterpaper and a4paper
papersize: a4paper

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 9pt

# Optional: Force one-column layout, default is two-column
#one_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
#numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography 
bibliography: bibliography

# Optional: Enable a 'Draft' watermark on the document
# watermark: true

# Customize footer, eg by referencing the vignette
# footer_contents: "YourPackage Vignette"

# Produce a pinp document
output: pinp::pinp

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{YourPackage-vignetteentry}
  %\VignetteKeywords{YourPackage, r, anotherkeyword}
  %\VignettePackage{YourPackage}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r Setup, include=FALSE, message=FALSE}
library("knitLatex")  #https://github.com/cran/knitLatex

f_normalize <- function(ds){
    f <- function(x){(x-min(x))/(max(x)-min(x))}   
    ds <- as.data.frame(sapply(ds, f))
    return(ds)
}

f_boxplot <- function(ds){
    library("ggplot2")
    ds.stack <- stack(ds)
    names(ds.stack)<-c("Values", "Features")
    p <- ggplot(ds.stack, aes(x = Features, y = Values, fill = Features)) + 
        geom_boxplot() + theme_minimal() + 
        theme(legend.position = "none", 
              axis.text.x = element_text(angle = 90, hjust = 1),
              axis.title.x=element_blank(),
              axis.title.y=element_blank())
    return(p)
}

f_cap <- function(ds){
    # Calculate IQR
    ds.iqr <- sapply(ds, IQR)
    ds.quantiles <- 
        as.data.frame(sapply(ds, 
                             function(x){quantile(x, c(0.05,0.25,0.75,0.95))}))
    
    # Calculate upper and lower limits
    ds.limits <- rbind(
        "lower" = ds.quantiles['25%',] - 1.5 * ds.iqr, 
        "upper" = ds.quantiles['75%',] + 1.5 * ds.iqr
    )
    # Replace outliers using capping method, by replacing values below 5th 
    # percentile and above 95th percentile by both lower and upper limits, 
    # respectively, where:
    # lower = Q1 - 1.5*IQR
    # upper = Q1 - 1.5*IQR
    for(n in names(ds)){
        ds[!is.na(ds[, n]) & 
                   ds[, n] < ds.limits["lower", n], n] <- 
            ds.quantiles["5%", n]
        ds[!is.na(ds[, n]) & 
                   ds[, n] > ds.limits["upper", n], n] <- 
            ds.quantiles["95%", n]
    }
    return(ds)
}

f_smote <- function(ds, class, formula, perc.over=100, perc.under=200, k=2){
    library("DMwR")
    set.seed(1)
    ds[,class] <- as.factor(ds[,class])
    ds <- SMOTE(formula, ds, 
                perc.over = perc.over,  
                perc.under = perc.under, 
                k=k)
    ds$quality <- as.numeric(as.character(ds[,class]))
    return(ds)
}

f_logistic <- function(ds.train, ds.test, cls, formula, 
                       family = binomial(link="logit"), 
                       predict.threshold=0.5){
    library("caret")
    library("pROC")
    
    set.seed(1)
    
    model <- glm(formula, ds.train, family=family)
    Y.test <- ds.test[, cls]
    Y_hat.test <- predict(model, ds.test, type="response")
    Y_hat.test[Y_hat.test >= predict.threshold] <- 1
    Y_hat.test[Y_hat.test < predict.threshold] <- 0
    cm <- confusionMatrix(data = as.factor(Y_hat.test), 
                          reference = as.factor(Y.test), 
                          positive='1')
    roc <- roc(Y.test, Y_hat.test, direction="<")
    
    r <- list()
    r$Y.test <- Y.test
    r$Y_hat.test <- Y_hat.test
    r$cm <- cm
    r$roc <- roc
    r$model <- model
    return(r)
}

f_logistic2 <- function(ds.train, ds.test, cls, formula, 
                       family = "binomial", 
                       alpha=0, lambda=1,
                       predict.threshold=0.5){
    library("glmnet")
    library("caret") 
    library("pROC")
    
    set.seed(1)
    
    Y.train <- ds.train[, cls]
    X.train <- model.matrix(formula, ds.train)
    Y.test <- ds.test[, cls]
    X.test <- model.matrix(formula, ds.test)
    
    model <- glmnet(X.train, Y.train, 
                    family=family, alpha=alpha, lambda = lambda)
    
    Y_hat.test <- as.numeric(predict(model, newx = X.test, type="response"))
    
    Y_hat.test[Y_hat.test >= predict.threshold] <- 1
    Y_hat.test[Y_hat.test < predict.threshold] <- 0
    cm <- confusionMatrix(data = as.factor(Y_hat.test), 
                          reference = as.factor(Y.test), 
                          positive='1')
    
    roc <- roc(Y.test, Y_hat.test, direction="<")
    
    r <- list()
    r$Y.test <- Y.test
    r$Y_hat.test <- Y_hat.test
    r$cm <- cm
    r$roc <- roc
    r$model <- model
    return(r)
}
```


## Introduction 

This report contains an analysis of **WineQuality** datasets, using a machine 
learning pipeline for exploring the datasets and proposing a logistic 
classifier model for identifying good and bad wines based on a few 
features.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.45\textwidth, height=2.5in]{pipeline} 
  \caption{Machine Learning Pipeline}\label{fig}
  \end{center}
\end{figure}

## Data Preparation

```{r DataLoading, echo=FALSE}
# Data Loading
fl.tra <- "wineQuality_train.data.txt"
fl.val <- "wineQuality_val.data.txt"
fl.tst <- "wineQuality_test.data.txt"
ds.tra <- read.csv(fl.tra)
ds.val <- read.csv(fl.val)
ds.tst <- read.csv(fl.tst)
```

There are **`r nrow(ds.tra)`** samples in the **training** dataset and 
**`r nrow(ds.val)`** samples in the **validation** dataset. 
The **validation** dataset represents 
**`r round(nrow(ds.val)/(nrow(ds.tra)+nrow(ds.val))*100, 2)`%**
of the total available data (both **training** and **validation** datasets).
The **testing** dataset contains **`r nrow(ds.tst)`** samples.
A few samples of the **training** dataset is shown in the 
**Table \ref{tab:TrainingDataset}**.

```{r TrainingDataset, echo=FALSE, results='asis'}
xTab(formatC(t(head(ds.tra, 6)), format='f', digits=2), 
     rows=TRUE, coldef = 'r|rrrrrr', 
     caption.bottom="\\label{tab:TrainingDataset}Training dataset")
```

```{r IncompleteCasesAnalysis, echo=FALSE}
ds.tra.nrows <- nrow(ds.tra[!complete.cases(ds.tra),])
ds.val.nrows <- nrow(ds.val[!complete.cases(ds.val),])
ds.tst.nrows <- nrow(ds.tst[!complete.cases(ds.tst),])
if (ds.tra.nrows==0 && ds.val.nrows==0 && ds.tst.nrows == 0){
    no_incomplete_cases <- TRUE
}
```

**Training**, **validation** and **test** datasets contains the following 
number of incomplete samples:

- **Training**: **`r ds.tra.nrows`** incomplete samples
- **Validation**: **`r ds.val.nrows`** incomplete samples
- **Testing**: **`r ds.val.nrows`** incomplete samples

```{r NoIncompleteCases, echo=FALSE, results='asis', eval=no_incomplete_cases}
cat("
As shown above, there are **no** incomplete cases in the **training**, 
**validation** and **testing** datasets.
")
```

```{r SummaryBeforeNormalization, echo=FALSE, results='asis'}
xTab(formatC(t(sapply(ds.tra, summary)), format='f', digits=2), 
      rows=TRUE, caption.bottom="\\label{tab:SummaryBeforeNormalization}Training dataset overview (without normalization)")
```

**Table \ref{tab:SummaryBeforeNormalization}** and 
**Table \ref{tab:SummaryAfterNormalization}**
present an overview of the **training** dataset before and after normalization.
For normalizing the datasets it was used the **min**-**max** normalization, as 
follows:

\begin{gather}
x'=\frac{x-min(x)}{max(x)-min(x)}
\end{gather}

```{r NormalizationForDistributionAnalysis, echo=FALSE, echo=FALSE}
ds.tra.norm <- ds.tra
ds.tra.norm[,1:11] <- f_normalize(ds.tra.norm[,1:11])

ds.val.norm <- ds.val
ds.val.norm[,1:11] <- f_normalize(ds.val.norm[,1:11])

ds.tst.norm <- ds.tst
ds.tst.norm[,1:11] <- f_normalize(ds.tst.norm[,1:11])
```

```{r SummaryAfterNormalization, echo=FALSE, results='asis'}
xTab(formatC(t(sapply(ds.tra.norm, summary)), format='f', digits=2), 
     rows=TRUE, caption.bottom="\\label{tab:SummaryAfterNormalization}Training dataset overview (normalized)")
```

The box plot in the **Figure \ref{fig:DistribuitionAnalysis}** gives a good 
overview of the data distribution of the **training** dataset before the 
removal of the outliers.
For the removal of the outliers, which are values that differ considerably from 
the majority of a set of data, different techniques are available. 
In this study, outlier removal was performed using the *capping* technique,by 
replacing values outside the $1.5*IQR$ limits, with 
the lower limit replaced by the **5th** percentile and the bigger limit replaced 
by the **95th** percentile.

```{r DistribuitionAnalysis, echo=FALSE, fig.cap="Data distribuition analysis"}
print(f_boxplot(ds.tra.norm[,1:11]))
```

The lower and upper limits are calculated by the equations:
\begin{gather}
IQR = Q3 - Q1 \\ 
Lower = Q1 - 1.5 * IQR \\ 
Upper = Q3 + 1.5 * IQR
\end{gather}

Once the limits are calculated, outliers below the **5th** percentile and above 
the **95th** percentile are replaced by both ***Lower*** and ***Upper*** limits, 
respectively.  

```{r OutliersCapping, echo=FALSE}
ds.tra[,1:11] <- f_cap(ds.tra[,1:11])
ds.val[,1:11] <- f_cap(ds.val[,1:11])
# ds.tst[,1:11] <- f_cap(ds.tst[,1:11])
```

```{r NormalizationAfterOutlierCleanup, echo=FALSE}
ds.tra.norm <- ds.tra
ds.tra.norm[,1:11] <- f_normalize(ds.tra.norm[,1:11])

ds.val.norm <- ds.val
ds.val.norm[,1:11] <- f_normalize(ds.val.norm[,1:11])

ds.tst.norm <- ds.tst
ds.tst.norm[,1:11] <- f_normalize(ds.tst.norm[,1:11])
```

After the **capping** of the outliers, the new data distribution is shown 
in **Figure \ref{fig:DistributionAnalysis2}**.

```{r DistributionAnalysis2, echo=FALSE, echo=FALSE, fig.cap="Data distribuition analysis after capping"}
print(f_boxplot(ds.tra.norm[,1:11]))
```

According to the quality, the datasets are balanced as follows:

```{r ImbalanceAnalysis, echo=FALSE}
ds.tra.gw.perc <- 
    round(nrow(ds.tra.norm[ds.tra.norm$quality==1,])*100/nrow(ds.tra.norm),2)
ds.val.gw.perc <- 
    round(nrow(ds.val.norm[ds.val$quality==1,])*100/nrow(ds.val.norm),2)
ds.tst.gw.perc <- 
    round(nrow(ds.tst.norm[ds.tst$quality==1,])*100/nrow(ds.tst.norm),2)
```

- **Tranining**: **`r paste0(ds.tra.gw.perc,"%")`** of good wine samples
- **Validation**: **`r paste0(ds.val.gw.perc,"%")`** of good wine samples
- **Test**: **`r paste0(ds.tst.gw.perc,"%")`** of good wine samples

## Model Training, Deployment and Evaluation

```{r Formulas, echo=FALSE, message=FALSE}
f1 <- formula(quality ~ .)

f2 <- formula( quality ~ fixed.acidity + volatile.acidity + citric.acid +
                   residual.sugar + chlorides + free.sulfur.dioxide +
                   total.sulfur.dioxide + density + pH + sulphates + alcohol + 
                   I(fixed.acidity ^ 2) + I(volatile.acidity ^ 2) + 
                   I(citric.acid ^ 2) + I(residual.sugar ^ 2) + I(chlorides ^ 2) + 
                   I(free.sulfur.dioxide ^ 2) + I(total.sulfur.dioxide ^ 2) + 
                   I(density ^ 2) + I(pH ^ 2) + I(sulphates ^ 2) + I(alcohol ^ 2))

f3 <- formula( quality ~ fixed.acidity + volatile.acidity + citric.acid +
                   residual.sugar + chlorides + free.sulfur.dioxide +
                   total.sulfur.dioxide + density + pH + sulphates + alcohol + 
                   I(fixed.acidity ^ 2) + I(volatile.acidity ^ 2) + 
                   I(citric.acid ^ 2) + I(residual.sugar ^ 2) + I(chlorides ^ 2) + 
                   I(free.sulfur.dioxide ^ 2) + I(total.sulfur.dioxide ^ 2) + 
                   I(density ^ 2) + I(pH ^ 2) + I(sulphates ^ 2) + I(alcohol ^ 2) +
                   I(fixed.acidity ^ 3) + I(volatile.acidity ^ 3) + 
                   I(citric.acid ^ 3) + I(residual.sugar ^ 3) + I(chlorides ^ 3) + 
                   I(free.sulfur.dioxide ^ 3) + I(total.sulfur.dioxide ^ 3) + 
                   I(density ^ 3) + I(pH ^ 3) + I(sulphates ^ 3) + I(alcohol ^ 3))

f4 <- formula( quality ~ fixed.acidity + volatile.acidity + citric.acid +
                   residual.sugar + chlorides + free.sulfur.dioxide +
                   total.sulfur.dioxide + density + pH + sulphates + alcohol + 
                   I(fixed.acidity ^ 2) + I(volatile.acidity ^ 2) + 
                   I(citric.acid ^ 2) + I(residual.sugar ^ 2) + I(chlorides ^ 2) + 
                   I(free.sulfur.dioxide ^ 2) + I(total.sulfur.dioxide ^ 2) + 
                   I(density ^ 2) + I(pH ^ 2) + I(sulphates ^ 2) + 
                   I(alcohol ^ 2) + I(fixed.acidity ^ 3) + I(volatile.acidity ^ 3) + 
                   I(citric.acid ^ 3) + I(residual.sugar ^ 3) + I(chlorides ^ 3) + 
                   I(free.sulfur.dioxide ^ 3) + I(total.sulfur.dioxide ^ 3) + 
                   I(density ^ 3) + I(pH ^ 3) + I(sulphates ^ 3) + I(alcohol ^ 3) +
                   I(fixed.acidity ^ 4) + I(volatile.acidity ^ 4) + 
                   I(citric.acid ^ 4) + I(residual.sugar ^ 4) + I(chlorides ^ 4) + 
                   I(free.sulfur.dioxide ^ 4) + I(total.sulfur.dioxide ^ 4) + 
                   I(density ^ 4) + I(pH ^ 4) + I(sulphates ^ 4) + I(alcohol ^ 4)   
               )

formulas <- c(f1, f2, f3)
```
For training the model, quadratic and cubic functions were created with 
the dataset features. Additionally, the ***SMOTE*** technique was used 
for dealing with the imbalanced datasets.

```{r "LogisticModel1", echo=FALSE, message=FALSE, warning=FALSE}
predict.threshold = 0.6

perc <- list(c(100,100), 
             c(100,150), 
             c(100,200),
             c(100,500))

quality_filter = "BACC"
min_quality_filter = 0.65

cross_analysis <- data.frame()
for(f_i in 1:length(formulas)){
    for(p_i in 1:length(perc)){
        ds.tra.norm.smote <- f_smote(ds.tra.norm, "quality",  
                                     formulas[[f_i]],
                                     perc.over = perc[[p_i]][1], 
                                     perc.under = perc[[p_i]][2], 
                                     k=2)
        ds.tra.norm.gw.perc <- 
            round(nrow(ds.tra.norm.smote[ds.tra.norm.smote$quality==1,])*100/
                      nrow(ds.tra.norm.smote),2)
        
        result <- f_logistic(ds.tra.norm.smote, ds.val.norm, "quality", 
                             formulas[[f_i]], 
                             predict.threshold = predict.threshold)
        roc <- result$roc
        cm <- result$cm
        t <- data.frame()
        t[1,"Formula"] <- f_i
        t[1,"BACC"] <- 
            formatC(cm$byClass["Balanced Accuracy"], format='f', digits=4)
        t[1,"F1"] <- formatC(cm$byClass["F1"], format='f', digits=4)
        t[1,"Good Wine Perc"] <- 
            formatC(ds.tra.norm.gw.perc, format='f', digits=2)
        
        if (t[1, quality_filter] >= min_quality_filter){
            cross_analysis <- rbind(cross_analysis, t[1,])
            assign(paste0("f", nrow(cross_analysis), "_model"),
                       result$model, envir = .GlobalEnv)
        }
    }
}
model.quality <- cross_analysis[, quality_filter]
best <- cross_analysis[model.quality==max(model.quality),]
model.var.name <- paste0("f", as.integer(as.numeric(rownames(best))), "_model")
model.best <- get(model.var.name)
formula.best <- get(paste0("f", best$Formula))
```

```{r CrossAnalysis, echo=FALSE, results='asis'}
rs <- rep("r", (ncol(cross_analysis) - 1))
rs <- paste0('r|', paste0(rs,collapse=""), collapse="")
xTab(cross_analysis, rows=FALSE, 
     coldef = rs,
     caption.bottom="\\label{tab:CrossAnalysis}Cross analysis matrix (without penalty terms)")
```

For fitting the model, the **`glm`** function was used for tuning the 
predefined functions, with the **training** dataset re-balancing during validation. 
The result of the analysis can be found in **Table \ref{tab:CrossAnalysis}**. 
The **confusion matrix** of the predicted wine quality against the 
actual classification, when using the **testing** dataset, can be found in the 
**Table \ref{tab:ConfusionMatrix}**. 
The performance during the tests is also shown by the **ROC** 
curve in **Figure \ref{fig:ROC1}**.

```{r TestVerification, echo=FALSE, message=FALSE, warning=FALSE}
Y <- ds.tst.norm[, "quality"]
Y_hat.pred <- predict(model.best, ds.tst.norm[, 1:11], type="response")
Y_hat <- Y_hat.pred
Y_hat[Y_hat >= 0.5] <- 1
Y_hat[Y_hat < 0.5] <- 0
cm <- confusionMatrix(data = as.factor(Y_hat), 
                      reference = as.factor(Y), 
                      positive='1')
roc <- roc(Y, Y_hat.pred, direction="<")
```

```{r ConfusionMatrix, echo=FALSE, results='asis'}
xTab(as.matrix(cm), rows=TRUE, 
     caption.bottom="\\label{tab:ConfusionMatrix}Confusion matrix (without penalty terms)")
```

```{r ROC1, echo=FALSE, fig.cap="ROC (without penalty terms)"}
plot(roc, col="blue", lwd=2)
```

```{r "LogisticModel2", echo=FALSE, message=FALSE, warning=FALSE}
predict.threshold = 0.6

perc <- list(c(100,100), 
             c(100,150), 
             c(100,200),
             c(100,500))

alpha <- c(0)
lambda <- c(0, 1e-07, 1e-06, 1e-05, 1e-04, 1e-03, 1e-02, 0.1, 1, 10)

quality_filter = "BACC"
min_quality_filter = 0.70

cross_analysis <- data.frame()
for(f_i in 1:length(formulas)){
    for(p_i in 1:length(perc)){
        for(l in lambda){
            for(a in alpha){
                ds.tra.norm.smote <- f_smote(ds.tra.norm, "quality",  
                                             formulas[[f_i]],
                                             perc.over = perc[[p_i]][1], 
                                             perc.under = perc[[p_i]][2], 
                                             k=2)
                ds.tra.norm.gw.perc <- 
                    round(nrow(ds.tra.norm.smote[ds.tra.norm.smote$quality==1,])*100/
                              nrow(ds.tra.norm.smote),2)
                
                result <- f_logistic2(ds.tra.norm.smote, ds.val.norm, "quality", 
                                      formulas[[f_i]], alpha = a, lambda=l,
                                      predict.threshold = predict.threshold)
                roc <- result$roc
                cm <- result$cm
                t <- data.frame()
                t[1,"Formula"] <- f_i
                t[1,"Alpha"] <- a
                t[1,"Lambda"] <- formatC(l, format='f', digits=1)
                t[1,"BACC"] <- 
                    formatC(cm$byClass["Balanced Accuracy"], format='f', digits=4)
                t[1,"F1"] <- formatC(cm$byClass["F1"], format='f', digits=4)
                t[1,"Good Wine Perc"] <- 
                    formatC(ds.tra.norm.gw.perc, format='f', digits=2)
                
                if (t[1,quality_filter] >= min_quality_filter){
                    cross_analysis <- rbind(cross_analysis, t[1,])
                    
                    assign(paste0("f", nrow(cross_analysis), "_model"),
                           result$model, envir = .GlobalEnv)
                }
            }
        }
    }
}
model.quality <- cross_analysis[,quality_filter]
best <- cross_analysis[model.quality==max(model.quality),]
model.var.name <- paste0("f", as.integer(as.numeric(rownames(best))), "_model")
model.best <- get(model.var.name)
formula.best <- get(paste0("f", best$Formula))
```

The function **`glmnet`** was also used for fitting the model, and during the 
evaluation of the quadratic and cubic functions, penalties were applied by 
changing the value of **`lambda`** parameter.
Similarly to the previous model fitting by **`glm`**, the **training** dataset 
was also re-balanced during the training. The result 
of that analysis can be found in **Table \ref{tab:CrossAnalysis2}**.
For avoiding spending space in this report, the **Table \ref{tab:CrossAnalysis2}**
does not bring all collected values in the training. Only the values with 
**Balanced Accuracy (BACC)** bigger than **70%** are shown. 
The **confusion matrix** for the predicted wine quality against the actual quality,
in the the **testing** dataset, can be found in **Table \ref{tab:ConfusionMatrix2}**. 
The performance during the tests of the classifier is also shown by the **ROC** 
curve in **Figure \ref{fig:ROC2}**.

```{r CrossAnalysis2, echo=FALSE, results='asis'}
rs <- rep("r", (ncol(cross_analysis) - 1))
rs <- paste0('r|', paste0(rs,collapse=""), collapse="")
xTab(cross_analysis, rows=FALSE, 
     coldef = rs,
     caption.bottom="\\label{tab:CrossAnalysis2}Cross analysis matrix (with penalty terms)")
```

```{r TestVerification2, echo=FALSE, message=FALSE, warning=FALSE}
model <- model.best
predict.threshold = 0.6
formula <- formula.best

Y.tst <- ds.tst.norm[, "quality"]
X.tst <- model.matrix(formula, ds.tst.norm)
    
Y_hat.prec <- predict(model, newx = X.tst, type="response")
Y_hat.tst <- Y_hat.prec
    
Y_hat.tst[Y_hat.tst >= predict.threshold] <- 1
Y_hat.tst[Y_hat.tst < predict.threshold] <- 0

cm <- confusionMatrix(data = as.factor(Y_hat.tst), 
                      reference = as.factor(Y.tst), 
                      positive='1')
roc <- roc(Y.tst, Y_hat.prec, direction="<")
```

```{r ConfusionMatrix2, echo=FALSE, results='asis'}
xTab(as.matrix(cm), rows=TRUE, 
     caption.bottom="\\label{tab:ConfusionMatrix2}Confision matrix (with penalty terms)")
```

```{r ROC2, echo=FALSE, fig.cap="ROC (with penalty terms)"}
plot(roc, col="blue", lwd=2)
```

## Final Conclusions

The logistic regression classifier with a generalized linear model with penalization, by means of the function **`glmnet`**, has big potential for bringing good results due to its fine-tuning parameters. Despite all the potentials of the **`glmnet`**, this study showed that the **`glm`** classifier performs well, giving almost similar results as the **`glmnet`** classifier. The dataset features were not enough for creating a good classifier, in both cases evaluated, and the addition of 3rd-degree components was necessary to increase the performance during training and validation. Additionally, the tuning of the classifier with a different proportion of wine classes showed that a better performance can be achieved by having more good wine samples in the training dataset.

