---
title: "House Pricing"
author: "Hirley Dayan Lourenço da Silva"
date: "4/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset Loading

```{r "Data Loading"}
f.training   <- "housePricing_trainSet.csv"
f.validation <- "housePricing_valSet.csv"

ds.training   <- read.csv(f.training)
ds.validation <- read.csv(f.validation)
```

## Data Analysis

There are **`r nrow(ds.training)`** samples in the traning dataset and 
**`r nrow(ds.validation)`** samples in the validation dataset. 

The validation dataset represents 
**`r round(nrow(ds.validation)/(nrow(ds.training)+nrow(ds.validation))*100, 2)`%**
of the total available data (both training and validation datasets).

- Training dataset **`r f.training`**:
```{r "Training Dataset"}
head(ds.training, 4)
```

- Validation dataset **`r f.validation`**:
```{r "Validation Dataset"}
head(ds.validation, 4)
```

The available features and its data types are listed following. There is only 
one categorical feature in both training and validation datasets that contains 
the house proximity to the ocean (`ocean_proximity`). In the case of linear 
regression, categorical features must not be used. In that sense, 
for this study, categorical features will be excluded.

- Training dataset **`r f.training`**:
```{r "Training Dataset Feature Classes"}
sapply(ds.training, class)
```

- Validation dataset **`r f.validation`**:
```{r "Validation Dataset Feature Classes"}
sapply(ds.validation, class)
```

Both training and validation datasets contains 
**`r nrow(ds.training[!complete.cases(ds.training),])`** and 
**`r nrow(ds.validation[!complete.cases(ds.validation),])`** inclomplete 
entries, respectivelly, and some samples are listed following. 
<font color="red">How to deal with missing data? Shall remove the lines or extrapolate?</font>

- Incomplete samples in the training dataset **`r f.training`**:
```{r "Incomplete Training Samples"}
head(ds.training[!complete.cases(ds.training),], 4)
```

- Incomplete samples in the Validation dataset **`r f.validation`**:
```{r "Incomplete Validation Samples"}
head(ds.validation[!complete.cases(ds.validation),], 4)
```

## Data Cleaning After Validation

Removal of categorical features:

- Training dataset **`r f.training`**:
```{r "Training Datasets Cleaning"}
ds.training <- ds.training[,-which(names(ds.training) %in% c("ocean_proximity"))]
head(ds.training, 4)
```


- Validation dataset **`r f.validation`**:
```{r "Validation Datasets Cleaning"}
ds.validation <- ds.validation[,-which(names(ds.validation) %in% c("ocean_proximity"))]
head(ds.validation, 4)
```

<font color="red">How to deal with missing data? Shall remove the lines or extrapolate?</font>


## Data Normalization

Data normalization using **z-score**:

- Training dataset **`r f.training`**:
```{r "Training Datasets Normalization"}
ds.training <-  as.data.frame(scale(ds.training, center = TRUE, scale = TRUE))
head(ds.training, 4)
```

- Validation dataset **`r f.validation`**:
```{r "Validation Datasets Normalization"}
ds.validation <-  as.data.frame(scale(ds.validation, center = TRUE, scale = TRUE))
head(ds.validation, 4)
```

## Linear Regression

```{r}
model <- paste(names(ds.training), collapse = '+')
model_formula <- as.formula(paste("median_house_value ~ ", model))
lm(model_formula, data = ds.training)
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
